---
output:
  html_document: default
  pdf_document: default
  word_document: default
---

### EXPLORATORY DATA ANALYSIS

Student Name: Clariza Look  



### Contents

1. **[Introduction](#Introduction) ** 

2. **[Description of Data](#DescriptionofData) ** 

3. **[Load Packages](#LoadPackages) ** 

4. **[Reading The Raw Data](#ReadingTheRawData) **

5. **[Data Understanding and Preparation](#DataUnderstandingandPreparation) **

6. **[Data Exploration](#DataExploration) **

8. **[References](#References) **


&nbsp;
&nbsp;
&nbsp;


### 1. Introduction {#Introduction}

This project is about conducting an exploratory data analysis using R of an airbnb dataset in Washington in 2015. It also includes in-depth analysis of data using visualizations and plots to extract insights. The EDA will help the analyst discover information about dataset's structure, determine critical variables, identify anomalies and evaluate assumptions.    

The methodology behind this process is to first understand the data by cleaning the data and converting the variables to the correct format. After that, we will extract the variables we wanted to do an analysis, then visualize and interpret results.


 

This data was made available by [Data World Washington 2015 Airbnb Dataset](https://data.world/codefordc/airbnb-washington-d-c-2015-10-03/workspace/file?filename=Listings+-+Detailed.csv). 

&nbsp;
&nbsp;
&nbsp;
 
### 2. Description of Data {#DescriptionofData}  

The file contains the scraped data captured by a web program with listing information from the airbnb website. This data may contain unformatted data points and many have duplicate entries. We'd want to clean and format the data prior to performing exploratory analysis, that will help better understand the available data and build some business context.  


**Business Understanding ** 

Before digging into data understanding and preparation, it is crucial that we first know our end goal or be aware of why this specific dataset is to be explored. There are many possible insights that we want to take from a few business questions to be asked but we will be focusing the questions to be answered during the process of the analysis.

**Problem Statements **  

Price

   + How are the listings distributed by location, by price, by type of property?
   + Which neighborhoods has the most expensive listings in DC?
   + What are the different types of listings in DC? Does their price vary by neighborhood?
   + Is there a relationships between price and other variables in the listing?




&nbsp;
&nbsp;
&nbsp;

### 3. Load Packages {#LoadPackages}  

```{r, echo = TRUE, message=FALSE}

library(plyr)
library(dplyr)
library(formattable)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(leaflet)
library(treemapify)


```


&nbsp;
&nbsp;
&nbsp;

### 4. Reading The Raw Data {#ReadingTheRawData}  

```{r, echo=TRUE}
airbnb_data <- read.csv('Listings.csv', header = TRUE, sep = ",")

kable(airbnb_data[1:7,1:28]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, font_size = 9) %>%
  scroll_box(width = "910px", height = "400px")

```



&nbsp;
&nbsp;
&nbsp;

### 5. Data Understanding and Preparation {#DataUnderstandingandPreparation} 

```{r, echo=TRUE}
dim(airbnb_data)

```

The listings.csv data has 3723 rows/observations (refer as the listings) and 96 columns (refer as attributes or features) per listing. 



```{r, echo = TRUE}

glimpse(airbnb_data)


```

+ These columns are of a character datatype (string format) with the dollar symbol '$' . If one of the these attributes will be used in the calculation, they need to be transformed into a numerical data. 


   + price          
   + weekly_price 
   + monthly_price  
   + security_deposit 
   + cleaning_fee      
   + extra_people  
   
+ Columns that should be categorical data are also not in the right format such as below:
   + neighbourhood
   + neighbourhood_cleansed
   + city
   + state
   + property_type 

If these attributes will be used in the analysis, they need to be transformed into a categorical data.

**Checking Missing Values **  

```{r, echo = TRUE}
colSums(sapply(airbnb_data, is.na))

```

   + Some data have null values. If one of the these attributes will be used in the data exploration, this needs to be addressed. 

**Remove Noise from Data **  

   + In this section we will filter the unnecessary columns from the dataset and therefore, we only choose relevant columns that we think we will be using in the analysis. Thus we will NOT include columns such as
      + id
      + listing_url
      + scrape_id
      + last_scraped
      + thumbnail_url
      + medium_url
      + picture_url
      + xl_picture_url
      + host_url 
      + host_thumbnail_url
      + host_picture_url
         
   + This will also include columns that has a large number of NULL values. 
         + license
         + square_feet
         + neighbourhood_group_cleansed
         
```{r, echo = TRUE}

#Remove unnecessary columns
new_airb_data <- select(airbnb_data, -id, -listing_url, -scrape_id, -last_scraped, -thumbnail_url,
                          -medium_url,-picture_url,-xl_picture_url,-host_url, -host_thumbnail_url, 
                          -host_picture_url, -license, -square_feet, -neighbourhood_group_cleansed)

dim(new_airb_data)

```      




&nbsp;
&nbsp;
&nbsp;

### 6. Data Exploration {#DataExploration}
   
   **Choosing the Relevant Columns For Analysis **  
   
   + Since the dataset is composed of properties in other states, we would like focus our analysis only for properties in DC.
      
```{r, echo=TRUE}

unique(new_airb_data[,"state"])

```

Load data that has properties only in Washington, DC
```{r, echo=TRUE}

#Filter data
dc_dataframe <- filter(new_airb_data,state == "DC")

## Count rows & columns
dim(dc_dataframe)

```
   Now we have a clean dataframe that will be ready for the next step which is doing the Price Analysis.  
   
   + **Chosen Columns and Their Descriptions **  
   + Based from the problem statements in above, we will be using these columns:  
      + "price": The price for a one-night stay. 
      + "property_type": refers to type of property
      + "room_type": refers to room type
      + "bedrooms": refers to number of bedrooms
      + "bed_type": refers to type of bed
      + "bathrooms": refers to number of bathrooms
      + "guests_included": refers to number of guests are included in the price 
      + "extra_people": refers to price for every extra person on top of "guests_included"
      + "accommodates": refers to number of people the property can accommodate
      + "minimum_nights": refers to the minimum number of nights guest could stay
      + "host_name": refers to name of host, names can be the same but these are actually different people
      + "host_id" refers to to ID of host, hosts can have 1 or more listings
      + "number_of_reviews" refers to number of people that gave reviews
      + "review_scores_rating" refers to rating from 30% (less positive) to 100% (positive)
      + "latitude" of listing
      + "longitude" of listing
      + "neighbourhood" The neighborhood location of the property  
         + For the "neighbourhood" data, some rows are missing values, so we will need more info from another column which is below  
            + "neighbourhood_cleansed:" The neighborhood location of the property with more detailed information   


```{r, echo = TRUE}

#Store columns in a new dataframe
airb_df <- dc_dataframe %>% select(price, property_type, room_type, bedrooms, bed_type, bathrooms, guests_included, extra_people, accommodates, neighbourhood, neighbourhood_cleansed, minimum_nights, host_name, host_id, number_of_reviews, review_scores_rating, latitude, longitude)

#check data in a formatted table
head(formattable(airb_df))


#Check data's datatype
glimpse(airb_df)


#Check NA or missing values
colSums(sapply(airb_df, is.na))
```


```{r}
#Check empty cells
colSums(airb_df == "")
```

#### **Insights From The Raw Data To Be Analyzed **
+ Looking at the data, we can see that 
   + It has 3,696 rows and 15 columns
   + "price" & "extra_people" are in a string form (e.g "$99.00"). They should be converted into float so we can use as a price format.  
   + categorical variables that needs to convert 
      + "property_type", "room_type", "bedrooms", "bed_type", bathrooms", "neighbourhood", "neighbourhood_cleansed", "host_name", "host_id", "review_scores_rating"
   
+ **NA values **: 
   + "bedrooms" has 21 NA values, "bathrooms" has 25 
   + "bathrooms" has a 0 value and it's impossible to have 0 bathroom in a property
      + We convert "Bedrooms" Na Values to 0 (assuming it's a studio-type property)
      + We convert "bathrooms" Na Values to 1 (assuming every property has at least 1 bathroom)
         + We convert "bathrooms" with 0 value to 1 (assuming every property has at least 1 bathroom)
   + "review_scores_rating" has 854, these can be replaced with 0 as it refers to 0 review  
   
+ **Empty Cells ** (different from NA Values but can impact the analysis):
   + "property_type" has 1 empty cell, we can place it with "Other" property_type
   + "neighbourhood" has 342 empty cells therefore we need extract details from                             "neighbourhood_cleansed" to fill in the empty cells
   

### Data Cleaning
```{r, echo=TRUE}

# Remove '$' sign in String from 'price' & 'extra_people'
# Replace new data in the "temp_price$price" column
temp_price <- gsub("\\$","", airb_df$price)
temp_extra_people <- gsub("\\$","", airb_df$extra_people)

# Because there are rows in the cell that has "," comma (e.g. 2,500.00)
# It automatically becomes NA when we convert them to float
# So to solve that problem, we have to remove "," in the columns
temp_price <- gsub("\\,","", temp_price)
temp_extra_people <- gsub("\\,","", temp_extra_people)

# Then convert number to float 
float_price <- as.double(temp_price)
float_extra_people <- as.double(temp_extra_people)

# Check datatype again to make sure there are no NA values
summary(float_price)
summary(float_extra_people)

#Now put back the cleaned data to "p_neighborhood_df$price"
airb_df$price <- float_price
airb_df$extra_people <- float_extra_people
```

#### A. Dealing with Empty cells in "neighbourhood" and "property_type" column
```{r, echo=TRUE}

## "neighbourhood" has 342 empty cells therefore we need extract details 
## from  "neighbourhood_cleansed" to fill in the empty cells
## Extract string after first comma and store it in new column "new_neighbourhood"
airb_df$new_neighbourhood <- sub(",.*", "", airb_df$neighbourhood_cleansed)

# Replace that empty "property_type" as "other"
airb_df$property_type[airb_df$property_type==''] <- "Other"

# Remove "neighbourhood","neighbourhood_cleansed" columns since we will not use them in
# the analysis as we created the "new_neighborhood" for analysis 
cleaned_airb_df <- subset(airb_df, select = -c(neighbourhood,neighbourhood_cleansed))

head(formattable(cleaned_airb_df))
str(cleaned_airb_df)


```

#### B. Dealing with the NA

```{r, echo=TRUE}

colSums(sapply(cleaned_airb_df, is.na))

##################################################################################
## We convert "bedrooms" Na Values to 0 (assuming it's a studio-type property)
## We convert "bathrooms" Na Values to 1 (assuming every property has at least 1 bathroom
## We convert "review_scores_rating" Na Values with 0 as it refers to 0 review
## We convert "bathrooms" with 0 value to 1 (assuming every property has at least 1 bathroom)
##################################################################################

cleaned_airb_df$bedrooms[is.na(cleaned_airb_df$bedrooms)] <- 0
cleaned_airb_df$bathrooms[is.na(cleaned_airb_df$bathrooms)] <- 1
cleaned_airb_df$review_scores_rating[is.na(cleaned_airb_df$review_scores_rating)] <- 0
cleaned_airb_df$bathrooms[cleaned_airb_df$bathrooms == 0] <- 1

## Check if NA values are gone
colSums(sapply(cleaned_airb_df, is.na))



```
#### B. Converting to categorical data

```{r, echo=TRUE}
# Convert  to categorical data
temp <- cleaned_airb_df %>% mutate_at(vars(room_type, bedrooms, bed_type, bathrooms, host_name, host_id, review_scores_rating, new_neighbourhood, property_type, bed_type),list(factor))
str(temp)

cleaned_airb_df <- temp


```


#### **Price Analysis **

##### **A. Checking Price Relationships with other variables** 

```{r, echo=TRUE}

pairs(~price+property_type+bedrooms+bed_type+minimum_nights+bathrooms+new_neighbourhood, data=cleaned_airb_df)

```

+ **Interpretation **  
   + There seems to have no clear patterns in the scatter plots that is interesting to do a linear correlation except for bedrooms and bathrooms variables.

##### **B. PRICE FREQUENCY**
```{r, echo=TRUE}

## Histogram of Price Frequency
densplot <- ggplot(data = cleaned_airb_df, mapping = aes(x = price))+
  geom_histogram(aes(y=..density..), bins=50, fill = "orange")+
  geom_density(color='dark green') +
  ggtitle("Property Rental Price Distribution in DC with Avg. Price")
  
   densplot + geom_vline(aes(xintercept=mean(price)),
           color="blue", linetype="dashed", size=1)

##Checking outliers in property price
boxplot(cleaned_airb_df$price)


```   
  
   + **Interpretation **  
   + The histogram depicts that the price is right skewed which means that the average property rental prices are greater than the median price in DC.  
   + Boxplot indicates that there are many outliers in prices that range from $400 to$2500. This could indicate that that these properties have house features which are making it more expensive. It can be more bedrooms,more amenities, prime location, or can accommodate more people     


&nbsp;
&nbsp;


##### **B. PRICE BY NEIGHBORHOOD**
```{r, echo=TRUE, message=FALSE}

## Group price by location and take average
p2 <- cleaned_airb_df %>% 
        group_by(new_neighbourhood) %>% 
        summarise(p_mean = mean(price),
                  p_median = median(price),
                  p_count= n()
                  )


ct <- ggplot(p2, mapping=aes(x=reorder(new_neighbourhood,p_count), y=p_count))+
  geom_col(fill="dark blue") +
  ggtitle("Properties By Neighborhood") + 
  xlab("Neighborhood") +
  ylab("Number of Listings") +
  coord_flip() +
  theme(axis.text.y=element_text(size=rel(0.8)))

mn <- ggplot(p2, mapping=aes(x=reorder(new_neighbourhood,p_count), y=p_mean))+
  geom_col(fill="#5B84B1FF") +
  ggtitle("Mean Price By Neighborhood") + 
  xlab("Neighborhood") +
  ylab("Mean Price") +
  coord_flip() +
  theme(axis.text.y=element_text(size=rel(0.8)))


md <-ggplot(p2, mapping=aes(x=reorder(new_neighbourhood,p_count), y=p_median))+
  geom_col(fill="dark green") +
  ggtitle("Median Price By Neighborhood") + 
  xlab("Neighborhood") +
  ylab("Median Price") +
  coord_flip() +
  theme(axis.text.y=element_text(size=rel(0.8)))




ct
mn
md


```   

   + **Interpretation **  
   + "Columbia Heights" has the most number of listings however the prices of the listings varies which affects the mean and median (price in Downtown has higher mean and median despite number of listings in the area are low)

&nbsp;
&nbsp;
&nbsp;

##### **C. PRICE BY PROPERTY TYPE **

```{r, echo=TRUE}

##Checking outliers in property price with mean points
prop_type_outliers <- ggplot(cleaned_airb_df, aes(x = price, y = property_type))

prop_type_outliers + geom_boxplot(notch = FALSE, fill = "lightgray") +
  ggtitle("Price By Property Type with Vs. Their Avg. Price") + 
  xlab("Price") +
  ylab("Property Type") +
  stat_summary(fun = mean, geom = "point",
               shape = 18, size = 2.5, color = "#FC4E07")

```


**Interpretation **  

   + We can see that "house" and "apartment" have many outliers than other types of property
   + The average price varies by property type

&nbsp;
&nbsp;
&nbsp;

##### **D. PRICE BY ROOM TYPE **

```{r, echo=TRUE}

ggplot(cleaned_airb_df, aes(x=price, fill=room_type)) +
  geom_density()


```


**Interpretation **  

   + According to the data, shared room has the most prices than any other room type
   
   


&nbsp;
&nbsp;
&nbsp;


##### **E. Room TypeBy Neighbourhood Treemap **

```{r, echo=TRUE}


tree_viz <- ggplot(cleaned_airb_df, aes(area = price, fill = room_type, subgroup = new_neighbourhood)) +
  geom_treemap(color = "red") +
  geom_treemap_subgroup_border(color = "white", size = 2) +
  geom_treemap_subgroup_text(place = "centre", grow = T, alpha = 1, colour =
                               "white", min.size = 1) +
  labs(title="Treemap: Room Type By Location") 
tree_viz

```

##### **E.1. Property Type By Neighbourhood Treemap **

```{r, echo=TRUE}


tree_viz1 <- ggplot(cleaned_airb_df, aes(area = price, fill = property_type, subgroup = new_neighbourhood)) +
  geom_treemap(color = "red") +
  geom_treemap_subgroup_border(color = "white", size = 3) +
  geom_treemap_subgroup_text(place = "centre", grow = T, alpha = 1, colour =
                               "white", min.size = 1) +
  labs(title="Treemap: Property Type By Location") 
tree_viz1

```

**Interpretation **  

   + The treemap depicts the rental prices among all room types and property types by location
   + This is to give an overview of prices by room_type and property types per location
   + Capitol Hill and Union Station seems to have high rental price for "Entire home/apt" in "E. Room TypeBy Neighbourhood Treemap"  while it also have high price in "house" property type in "E.1. Property Type By Neighbourhood Treemap"
   + In "E.1. Property Type By Neighbourhood Treemap" West End", "Downtown" and "Kalorama Heights" has the most apartments compare to other property types in their respective areas. 
   
   


&nbsp;
&nbsp;
&nbsp;

##### **F. Price By Number of Bedrooms **

```{r, echo=TRUE}


violin_bdrm <- ggplot(cleaned_airb_df, aes(x=price, y=bedrooms, color=bedrooms)) + 
  geom_violin(trim=FALSE) +
   coord_flip() +
   stat_summary(fun=mean, geom="point", size=2, color="red")

violin_bdrm 




```


**Interpretation **  

   + The violin plot represents the distribution of price per number of bedrooms related to their outlier prices
   + It shows there are many properties with 0 bedrooms (assuming it is a studio type apartment) and rental prices are just below $100.
   + It shows that the 5 bedrooms has a wide range of min to max prices
  


   

&nbsp;
&nbsp;
&nbsp;




##### **F. Number of Bedrooms by Property Type**

```{r, echo=TRUE}


ggplot(cleaned_airb_df, aes(x=bedrooms, fill=property_type)) +
  geom_bar() 


```


**Interpretation **  

   + The stack bar plot represents the number of bedrooms by property  
   + The data shows that most of the "Apartments" listed are having 1 bedroom  
   + It also shows that all "boats" listings have 1 bedroom
   
   
&nbsp;
&nbsp;
&nbsp;


##### **H.  Property Type by BedType**

```{r, echo=TRUE}


ggplot(cleaned_airb_df, aes(x=property_type)) +
  geom_bar() +
   coord_flip() +
  theme(axis.text.y=element_text(size=rel(0.8)))



```


**Interpretation **  

   + Many of the property types are using "read beds" than any other bedtype
   + The data shows that the least number of bed type being used is couch

&nbsp;
&nbsp;
&nbsp;


##### **I. Number Accommodates by Property Type **

```{r, echo=TRUE}

pss <- ggplot(cleaned_airb_df, aes(property_type, accommodates))
pss + geom_point(colour="blue") +
   coord_flip() +
   ylab("Accommodates") +
   xlab("Property Type") +
   ggtitle("Number Accommodates by Property Type") +
  theme(axis.text.y=element_text(size=rel(0.8))) 
   
```


**Interpretation **  

   + House has the most flexible number of accommodations among all the property types
   + Dorm seems to accept up to 6 people in one stay while boat can take up to 4 people

&nbsp;
&nbsp;
&nbsp;

##### **I. Number of Reviews by Property Type **

```{r, echo=TRUE}

pss <- ggplot(cleaned_airb_df, aes(property_type, number_of_reviews))
pss + geom_point(colour="dark green") +
   coord_flip() +
   ylab("Reviews") +
   xlab("Property Type") +
   ggtitle("Number Reviews by Property Type") +
  theme(axis.text.y=element_text(size=rel(0.8))) 
   
```


**Interpretation **  

   + Reviews indicate the number of people who have stayed and left a review
   + House and Apartment has the most number of reviews which could possibly mean that during this time of the data was captured, House and Apartment property types are the most stayed property type.

&nbsp;
&nbsp;
&nbsp;


##### **J. Listing Map **

```{r, echo=TRUE}



expensive_listings <- cleaned_airb_df %>% 
  filter(cleaned_airb_df$price >= 500) 

inexpensive_listings <- cleaned_airb_df %>% 
  filter(cleaned_airb_df$price < 500) 

listing_map <- leaflet() %>% 
   addTiles() %>%
   setView(lng = -77.046331, lat = 38.909702 , zoom = 12) %>%
   addCircleMarkers( lng = inexpensive_listings$longitude, 
                     lat = inexpensive_listings$latitude,
                     radius = 2, 
                     stroke = FALSE,
                     color = "blue",
                     fillOpacity = 0.5, 
                     group = "Inexpensive Properties"
                     )  %>%
   addCircleMarkers( lng = expensive_listings$longitude, 
                     lat = expensive_listings$latitude,
                     radius = 4, 
                     stroke = FALSE,
                     color = "red",
                     fillOpacity = 0.5, 
                     group = "Expensive Properties"
                     )


listing_map 
   
```


**Interpretation **  

   + The listings map above represents the high price rental properties (price above $500) vs. low price (price below $500) properties
   + Looking at the map alone does not clearly specify patterns of high price based on just the location having features such as "prime spot", or "near the park", "near the airport" , "near the river" , etc. 

&nbsp;
&nbsp;
&nbsp;

#### 7. Summary & Conclusions {#SummaryConclusions}


The data analysis done above a great stat to give insights about pricing by different variables. 

**Problem Statements **  

+ How are the listings distributed by location, by price, by type of property?
   + They are distributed around Washington DC and the price varies by property type, number of bedrooms and perhaps some other variables that is not covered by our dataset
+ Which neighborhoods has the most expensive listings in DC?
   + In the "Listing Map" map section, it shows the distribution of high priced listing compare to low priced listings by map level. However in the "B. PRICE BY NEIGHBORHOOD" section, it states that based on median price, Georgetown neighbourhood has the highest price. 
+ What are the different types of listings in DC? Does their price vary by neighborhood?
   + These includes Townhouse, Loft, House, Dorm, Condominium, Cabin, Bungalow, Boat, Bed & Breakfast, Apartment
   + In the "E. Room Type Treemap" section, it shows the prices by room type per location which gives us a rough idea of prices by neighborhood
+ Is there a relationships between price and other variables in the listing?
   + This part needs further analysis and perhaps would need more variables in the dataset or perhaps analyse the "description" column and extract details about property features. 

However we can also do further analysis on:

 + Price with date and time to determine demand/supply of house price rentals in the location
 + Revenue with all variables (similar to pricing analysis) to determine which location is best to invest an airbnb property
 
 
 
&nbsp;
&nbsp;
&nbsp;

#### 8. References {#References}




+ http://www.sthda.com/english/articles/32-r-graphics-essentials/132-plot-grouped-data-box-plot-bar-plot-and-more/
+ https://www.kaggle.com/chrisbow/e-commerce-eda-and-segmentation-with-r  
+ https://rkabacoff.github.io/datavis/Bivariate.html 
+ https://www.kaggle.com/notaapple/detailed-exploratory-data-analysis-using-r  
+ https://analyticsindiamag.com/tutorial-get-started-with-exploratory-data-analysis-and-data-preprocessing/
+ https://learn.r-journalism.com/en/mapping/leaflet_maps/leaflet/ 


&nbsp;
&nbsp;
&nbsp;



